---
title: "08_data_cleaning"
output: html_document
---

For data wrangling, we normally do data cleaning, standardize, manipulate, and visualize data.

## Data Cleaning

Useful `R` packages for data cleaning:
- `Here`: Makes referencing files easier
- `skimr`: Makes summarizing data easier and let us skim through it faster
- `Janitor`: Simplify data cleaning tasks

```{r}
renv::install('here')
```
```{r}
renv::install('skimr')
```
```{r}
renv::install('janitor')
```
## Import necessary libraries
```{r}
library('here')
library('skimr')
library('janitor')
library('dplyr')
```

## Load Palmer Penguins dataset
```{r}
renv::install('palmerpenguins')
library('palmerpenguins')
```
## Explore dataset
```{r}
skim_without_charts(penguins)
```
```{r}
glimpse(penguins)
```

```{r}
head(penguins)
tail(penguins)
```
To select only species column:
```{r}
penguin_species <- penguins %>% 
                      select(species)
penguin_species
```

To omit only `species` column and shows the rest:
```{r}
penguin_omit_species <- penguins %>% 
                          select(-species)
penguin_omit_species
```

To rename column:

```{r}
penguins %>% 
  rename(island_new = island)
```

We can change the column name either to capital or small letters:

```{r}
rename_with(penguins, toupper)
```

```{r}
rename_with(penguins, tolower)
```

To ensure there's only characters, numbers, and underscores in column names:

```{r}
clean_names(penguins)
```

