---
title: "11_transforming_data"
output: html_document
---

## Manually create data frame

```{r}
id <- c(1:10)

name <- c("John Mendes", "Rob Stewart", "Rachel Abrahamson", "Christy Hickman", "Johnson Harper", "Candace Miller", "Carlson Landy", "Pansy Jordan", "Darius Berry", "Claudia Garcia")

job_title <- c("Professional", "Programmer", "Management", "Clerical", "Developer", "Programmer", "Management", "Clerical", "Developer", "Programmer")

employee <- data.frame(id, name, job_title)
```

## Import necessary library
```{r}
library(tidyr)
```

## Important transforming functions

There are 3 important functions in `R` for data transformation:
- `separate()`
- `unite()`
- `mutate()`


```{r}
print(employee)
```

### Separate name into fore name and last name

```{r}
separate(employee, name, into=c('first_name', 'last_name'), sep = ' ')
```

### Unite columns into single column

```{r}
first_name <- c("John", "Rob", "Rachel", "Christy", "Johnson", "Candace", "Carlson", "Pansy", "Darius", "Claudia")

last_name <- c("Mendes", "Stewart", "Abrahamson", "Hickman", "Harper", "Miller", "Landy", "Jordan", "Berry", "Garcia")

job_title <- c("Professional", "Programmer", "Management", "Clerical", "Developer", "Programmer", "Management", "Clerical", "Developer", "Programmer")

employee <- data.frame(id, first_name, last_name, job_title)

print(employee)
```
```{r}
unite(employee, 'name', first_name, last_name, sep = ' ')
```

### Add columns with Calculation using `mutate()`

```{r}
View(penguins)
```

E.g. body_mass is measured in gram. Let's say we need another column of body mass in kg:

```{r}
penguins %>% 
  mutate(body_mass_kg = body_mass_g/1000, flipper_length_m = flipper_length_mm/1000)
```

