---
title: "12_bias_function"
output: html_document
---

Bias function denotes by the average amount that the actual outcome is greater than the predicted outcome.

- **Interpretation of Results:**
    - **Close to Zero:** Indicates an **unbiased** model or prediction.
    - **High Result (positive or negative):** Suggests that your data or model might be **biased**.
        - A positive bias means actual values are consistently higher than predicted.
        - A negative bias means actual values are consistently lower than predicted (predictions are too high).
- **Benefit:** Provides actionable insights to identify and address systemic issues causing inaccurate predictions.

## Import bias library

```{r}
renv::install('SimDesign')
```
```{r}
library(SimDesign)
```

## Compare actual temperature vs forecast

```{r}
actual_temp <- c(32.3, 35.6, 33.3, 36.6, 35.6, 36.7)
forecast_temp <- c(32.0, 34.5, 32.2, 35.6, 34.4, 35.5)

bias(actual_temp, forecast_temp)
```

## Compare sales result bias

```{r}
actual_sales <- c(153, 314, 234, 524, 422, 243)
predicted_sales <- c(233, 432, 524, 252, 243, 423)

bias(actual_sales, predicted_sales)
```

