---
title: "16_aesthetics"
output: html_document
---

## Import necessary libraries

```{r}
library(ggplot2)
library(palmerpenguins)
```

## Plotting

```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```

## Manipulate the shape, size, and color for datapoints

```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, color=species))
```
 ### Shape
 
```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, shape=species))
```
### Color & Shape


```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, color=species, shape=species))
```

### Using Alpha for dominant datapoints for high density datasets

This affect transparencey or color intensity of the aesthetics.
```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, alpha=species))
```
### Generalize color without mapping


```{r}
ggplot(data=penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g), color="green")
```

## Geom functions

```{r}
ggplot(data=penguins) +
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
ggplot(data=penguins) +
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```
### Linetype


```{r}
ggplot(data=penguins) +
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g, linetype = species))
```
## Jitter


```{r}
ggplot(data=penguins) +
  geom_jitter(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```

## Bar charts

```{r}
ggplot(data=diamonds) +
  geom_bar(mapping = aes(x=cut))
```
### Manipulate bar colors
```{r}
ggplot(data=diamonds) +
  geom_bar(mapping = aes(x=cut, color=cut))
```
To get fill colored bar instead of just the outline colored:
```{r}
ggplot(data=diamonds) +
  geom_bar(mapping = aes(x=cut, fill=cut))
```
### stacked bar chart

```{r}
ggplot(data=diamonds) +
  geom_bar(mapping = aes(x=cut, fill=clarity))
```

## Smoothing

Smoothing helps data professionals reveal trends. When scatterplots alone lack clarity, smoothing adds a trend line, making underlying patterns in the data easier to spot for casual observers. ggplot2 offers two smoothing methods: Loess is best for plots with fewer than 1,000 points, it creates a flexible, local smoother. Gam is ideal for larger datasets because it uses a more robust model for general trends. Smoothing enhances data communication, adding a visual cue to highlight trends so data visualizations become clearer and more impactful for audiences.









