---
title: "19_annotations"
output: html_document
---

## Import important libraries

```{r}
library(ggplot2)
library(palmerpenguins)
```
## Labels and Annotations

Labels add text outside the grid box.

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title = 'Palmer Penguins: Body Mass vs. Flipper Length', subtitle = 'Sample of Three Penguins Species',
       caption = 'Data collected by Dr. Kristen Gorman in 2007-2009')
```

Annotations add text inside the grid box:
```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title = 'Palmer Penguins: Body Mass vs. Flipper Length', subtitle = 'Sample of Three Penguins Species',
       caption = 'Data collected by Dr. Kristen Gorman in 2007-2009') +
  annotate('text', x=220, y=3500, label='The Gentoos are the largest', color='medium blue',
           fontface='bold', size=4.5, angle=25)
```

## Using variable for base plot

```{r}
p <- ggplot(data = penguins) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title = 'Palmer Penguins: Body Mass vs. Flipper Length', subtitle = 'Sample of Three Penguins Species',
       caption = 'Data collected by Dr. Kristen Gorman in 2007-2009')

p + annotate('text', x=220, y=3500, label='The Gentoos are the largest', color='medium blue',
             fontface='bold', size=4.5, angle=25)
```

## Savings viz

1) Using Export feature of RStudio or Posit Cloud or `Save image as..` option in R Markdown.
2) Using ggplot `ggsave()` function. By default, it saves the last image generated in notebook or R console.
3) Using graphic devices. A graphics device allows a plot to appear on your computer. Examples include:

  - A window on your computer (screen device)
  
  - A PDF, PNG, or JPEG file (file device)
  
  - An SVG, or scalable vector graphics file (file device)
  

### Using ggplot

```{r}
ggsave('three_penguin_species.png')
```

### Using graphic devices

#### `dev.off()`

For png:
```{r}
png(file = "exampleplot.png", bg = "transparent")
plot(1:10)
rect(1, 5, 3, 7, col = "white")
dev.off()
```
For pdf:
```{r}
pdf(file = "/Users/username/Desktop/example.pdf",    
       width = 4,     
       height = 4) 
plot(x = 1:10,     
        y = 1:10)
abline(v = 0)
text(x = 0, y = 1, labels = "Random text")
dev.off()
```



